---
title: "etoAQS"
output: JSON of all AQS data
---
created: 10/10/2022

# setup
## libraries
- RAQSAPI for AQS query [GitHub (get started)](https://github.com/USEPA/RAQSAPI); [user manual](https://cran.r-project.org/web/packages/RAQSAPI/RAQSAPI.pdf)
- keyring for credentials
- lubridate for easier datetime handling
- writexl for storing table
- data.table for making tables
  
```{r}
library(RAQSAPI) 
library(keyring)
library(lubridate)
library(writexl)
library(data.table)
```

set credentials; API key is dunbird69

```{r} 
#uncomment below if first time user; only need to do this once
#aqs_sign_up("emei3@gatech.edu") #select email to sign up
```

```{r}
datamartAPI_user <- "emei3@gatech.edu" #change to match your email
server <- "AQSDatamart"
keyring::key_set(service = server,
                 username = datamartAPI_user) #enter API key when prompted
```

get access to API; if your credentials are wrong, you will get no indication. However, your other functions will not work. 

```{r}
aqs_credentials(username = datamartAPI_user,
                key = key_get(service = server,
                              username = datamartAPI_user
                              )
                )
```

# Query

- eto pollutant code is 43601
- beg date was chosen as before all monitoring efforts began. End date is selected as today to obtain most recent data
- all 50 states were chosen to be queried for the inital run
  - statesWithData indicates that only 17 states have data
  - in subsequent runs, only data from these states are pulled

```{r}
pollutant <- "43601" #ethylene oxide
begDate <- as.Date("2018-01-01") #beginning year of eto monitoring
endDate <- as.Date(today()) #today
#state <- 1:50 #FIPS codes for all states; reduce number if you know certain states are null
state <- c(4, 6, 8, 12, 13, 17, 21, 25, 26, 29, 
           34, 36, 40, 42, 44, 45, 49, 53, 55) #determined from prior run

## preallocate data
statesWithData <- c()
allData <- data.table()


## query by state (if n years and m states, will take n*m*(5 seconds) to run)
for (x in state) {
  temp = aqs_sampledata_by_state(parameter = pollutant,
                        bdate = begDate,
                        edate = endDate,
                        stateFIPS = sprintf("%02d", x)) #query data
  
  allData <- rbind(allData, temp) #add queried data to list
  
  if (nrow(temp) != 0) {
    statesWithData <- append(statesWithData, x)
  }
}

```


write the table to a file
```{r}
fp <- "../Data/" #desired file path

write_xlsx(allData, paste(fp, "etoAQS.xlsx", sep = ""))
```

# temporary fix for missing states
- in initial search for AQS data, only FIPS codes 1:50 were queried, which misses a few states

```{r}
# pollutant <- "43601" #ethylene oxide
# begDate <- as.Date("2018-01-01") #beginning year of eto monitoring
# endDate <- as.Date(today()) #today
# #state <- 1:50 #FIPS codes for all states; reduce number if you know certain states are null
# state <- 51:56 #missing FIPS codes, only 53 and 55 have data
# 
# ## preallocate data
# statesWithDataAddendum <- c()
# allDataAddendum <- data.table()
# 
# 
# ## query by state (if n years and m states, will take n*m*(5 seconds) to run)
# for (x in state) {
#   temp = aqs_sampledata_by_state(parameter = pollutant,
#                         bdate = begDate,
#                         edate = endDate,
#                         stateFIPS = sprintf("%02d", x)) #query data
#   
#   allDataAddendum <- rbind(allDataAddendum, temp) #add queried data to list
#   
#   if (nrow(temp) != 0) {
#     statesWithDataAddendum <- append(statesWithDataAddendum, x)
#   }
# }
# 
# allData <- rbind(allData, allDataAddendum) #add new queried data to list
```

add data to new file

```{r}

# fp <- "../Data/" #desired file path
# 
# write_xlsx(allDataAddendum, paste(fp, "etoAQSAddendum.xlsx", sep = ""))

```
